Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-06-10 response-to-reviews.Rmd

l.744 'Because canopy trees dominate these fluxes (Fig. 5), their responses will strongly influence the whole-ecosystem response'

l.898 'disproportionate role of large trees in ecosystem dynamics'

So perhaps the authors can greatly soften their ecosystem level assessment. To their credit, and what truly should be emphasised, is the further research needed on ecosystem function (l. 757).

**We have reworded the abstract sentence to place less emphasis on ecosystems: "Using published research and new analyses, we evaluate how microclimate and leaf temperatures, traits, and gas exchange vary vertically in forests, shaping tree and ecosystem ecology." This wording parallels the section headers in the main manuscript. While we agree that there is little information on how ecosystem function varies across vertical profiles, this was included within the scope of our review, and we dedicate a 450-word subsection (sec. 5.2) plus a figure to this theme.**

### 2022-06-10 vertical-thermal-review.Rmd
response on ecosystem in abstract (R2)
@NidhiVinod , FYI. I believe I've now dealt with all of the most difficult reviewer comments. There are ~5 wording changes requested by R2 and a few small technical corrections from R1 remaining. I'm stopping for now and expect that you can easily handle the rest, but please ping me if you need help.

Rising temperatures are influencing forests on many scales, with potentially strong variation vertically across forest strata. 
Using published research and new analyses, we evaluate how microclimate and leaf temperatures, traits, and gas exchange vary vertically in forests, shaping tree ecology and ecosystem function. 
In closed-canopy forests, upper-canopy leaves are exposed to the highest solar radiation and evaporative demand, which can elevate leaf temperature (*T~leaf~*), particularly when transpirational cooling is curtailed by limited stomatal conductance.
However, foliar traits also vary across height or light gradients, partially mitigating and protecting against the elevation of upper-canopy *T~leaf~*.
Leaf metabolism generally increases with height across the vertical gradient, yet differences in thermal sensitivity across the gradient appear modest.
Scaling from leaves to trees, canopy trees have higher absolute metabolic capacity and growth, yet are more vulnerable to drought and damaging *T~leaf~* than their smaller counterparts, particularly under climate change. 
In contrast, understory trees experience fewer extreme high *T~leaf~*'s but have fewer cooling mechanisms and thus may be strongly impacted by warming under some conditions, particularly when exposed to a harsher microenvironment through canopy disturbance.
As the climate changes, integrating the patterns and mechanisms reviewed here into models will be critical to forecasting forest-climate feedbacks.

>>>>

Rising temperatures are influencing forests on many scales, with potentially strong variation vertically across forest strata. 
Using published research and new analyses, we evaluate how microclimate and leaf temperatures, traits, and gas exchange vary vertically in forests, shaping tree and ecosystem ecology. 
In closed-canopy forests, upper-canopy leaves are exposed to the highest solar radiation and evaporative demand, which can elevate leaf temperature (*T~leaf~*), particularly when transpirational cooling is curtailed by limited stomatal conductance.
However, foliar traits also vary across height or light gradients, partially mitigating and protecting against the elevation of upper-canopy *T~leaf~*.
Leaf metabolism generally increases with height across the vertical gradient, yet differences in thermal sensitivity across the gradient appear modest.
Scaling from leaves to trees, canopy trees have higher absolute metabolic capacity and growth, yet are more vulnerable to drought and damaging *T~leaf~* than their smaller counterparts, particularly under climate change. 
In contrast, understory trees experience fewer extreme high *T~leaf~*'s but have fewer cooling mechanisms and thus may be strongly impacted by warming under some conditions, particularly when exposed to a harsher microenvironment through canopy disturbance.
As the climate changes, integrating the patterns and mechanisms reviewed here into models will be critical to forecasting forest-climate feedbacks.
